---
title: "Entrega-1"
output: html_document
date: '2023-02-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## R Markdown

### Transformació de les dades

Obtenim les dades:

```{r dataset, include=TRUE}

df <- read.csv2("./bank-additional.csv")
```

Especifiquem aquelles dades que hem convertit i les transformacions que hem aplicat en elles.

#### Age:

Quatre possibles categories: [0,25] [26-45] [46-65] [66-inf]

```{r}
df$age <- cut(df$age, 
              breaks = c(0, 25, 45, 65, max(df$age)),
              labels = c("A", "B", "C", "D"))
```

#### Job:

admin. blue-collar entrepreneur housemaid management retired self-employed services student technician unemployed unknown

```{r}
df %>% mutate(job = ifelse(job == "retired" | job == "unemployed" | job == "student" | job == "housemaid", "unemployed", job))
df %>% mutate(job = ifelse(job == "entrepeneur", "self-employed", job))
```

#### Marital:

divorced married single unknown

```{r}

```

#### Education:

```{r}
df %>% mutate(education = ifelse(education, education == "basic.4y" | education == "basic.6y" | education == "basic.9y", "basic", education))
df %>% mutate(education = ifelse(education, education == "unknown", NA , education))
```

#### Default:

```{r}

```

#### Housing:

```{r}

```

#### Loan:

```{r}

```

#### Contact:

```{r}

```

#### Month

```{r}

```

#### Day of week:

```{r}

```

#### Duration:

```{r}
df <- cut(df$duration,
          breaks = c(0, 30, 300, max(df$duration)),
          labels = c("Baixa", "Mitja", "Alta"))
```

#### Campaign:

```{r}

```

#### Pdays:

```{r}
df %>% mutate(pdays = ifelse(pdays, pdays == 999, NA , pdays))
```

#### Previous:

```{r}

```

#### Poutcome:

```{r}

```

#### Employment variation rate

```{r}

```

#### Consumer price index

```{r}

```

#### Consumer confidence index

```{r}

```

#### Euribor 3 month rate

```{r}

```

#### Number of employees

```{r}

```
